---
output:
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	echo=FALSE,
	dpi=300
	)
```

# Peabody Invertebrate Zoology Collections Statistics

This document describes a variety of statistics about the [Peabody Invertebrate Zoology Collections](http://peabody.yale.edu/collections/invertebrate-zoology).

The data were obtained from Daniel Drew on January 8, 2018.

```{r}
library( tidyverse )
library( magrittr )
library( knitr )
library( plotly )


D = read_tsv( "IZReport_20180108.tsv" )

nrow(D)
names(D)





```

```{r asis=TRUE }

summary_stats = data.frame(Description=character(), N=numeric())
summary_stats %<>% rbind(list( "All Records", nrow(D) ), stringsAsFactors=FALSE )
summary_stats %<>% rbind(list( "Lat and Lon", sum( 
	!is.na(D$`Latitude (dec)`) & !is.na(D$`Longitude (dec)`), 
	stringsAsFactors=FALSE
	) ) )

names(summary_stats) = c("Description", "N")

kable(summary_stats)

```


```{r asis=TRUE}
phylum_list = table(D$Phylum)
phylum_frame = data.frame(Taxon=names(phylum_list), N=phylum_list)
kable( phylum_frame )
```


```{r}
ggplot(D) + 
	geom_freqpoly(aes(x=`Collecting Year`, y=..count..)) +
	xlim(1800, 2017)
```
```{r}

collectors = c(table(D$`Collector/Expedition`))
collectors = collectors[order(collectors, decreasing = TRUE)]

collectors_frame = data.frame(Name=names(collectors), N=collectors, stringsAsFactors = FALSE)
rownames(collectors_frame) = NULL
kable( collectors_frame[1:100,] )
```


## Locations

```{r}

D_ll = D[ (!is.na(D$`Latitude (dec)`) & !is.na(D$`Longitude (dec)`) ), ]


mapWorld <- borders("world", colour="gray50", fill="gray50") # create a layer of borders

mp = 
	ggplot( D_ll ) +   
	mapWorld + 
	geom_point(
		aes(x=`Longitude (dec)`, y=`Latitude (dec)`, col=`Collecting Year`, label1=`Catalog Number` ), alpha = 0.2, size=1
		) +
	scale_color_gradient(low="blue", high="red")

mp
#ggplotly(mp)


```

## Arianna queries

### Most prolific collectors

Most collections are from the US Fish Commission cruises, so I'll separate these out.

```{r}

fish_string = "^U.S. Fish Comm"

collectors_focal = 
	collectors_frame %>%
	filter( str_detect(Name, fish_string) )

D_ll %>%
	filter( `Collector/Expedition` %in% collectors_focal$Name[1:10] ) %>%
	ggplot( ) +  
		mapWorld +
		geom_point(aes(x=`Longitude (dec)`, y=`Latitude (dec)`, col=`Collector/Expedition` ), alpha = 0.2, size=1) +
	  guides(colour = guide_legend(override.aes = list(alpha = 1)))

```

Here are the non US Fish Commission cruises collectors.


```{r}

collectors_focal = 
	collectors_frame %>%
	filter( ! str_detect(Name, fish_string) )

map_not_fish =
D_ll %>%
	filter( `Collector/Expedition` %in% collectors_focal$Name[1:10] ) %>%
	ggplot( ) +  
		mapWorld +
		geom_point(aes(x=`Longitude (dec)`, y=`Latitude (dec)`, colour=str_wrap (`Collector/Expedition`,20) ), alpha = 0.2, size=1) +
  labs(colour="Collecter/Expedition") +
	  guides(colour = guide_legend(override.aes = list(alpha = 1)))

map_not_fish

```

## Lazo-Wasem_string

Eric has been a very prolific collector, however his name as a collector varies within the collection due to differnet collection events/other co-collectors 

Below I have seperated these out.

```{r}

LazoWasem_string = "^Lazo-Wasem"
 
collectors_focal = 
    collectors_frame %>%
    filter( str_detect(Name, LazoWasem_string) )

D_ll %>%
  filter( `Collector/Expedition` %in% collectors_focal$Name[1:18] ) %>%
  ggplot( ) +  
  mapWorld +
  geom_point(aes(x=`Longitude (dec)`, y=`Latitude (dec)`, colour=str_wrap (`Collector/Expedition`,40) ), alpha = 0.2, size=1) +
  labs(colour="Collecter/Expedition") +
  guides(colour = guide_legend(override.aes = list(alpha = 1)))
  

```




## Sub regions

```{r}


mapPacific <- borders("world", colour="gray50", fill="gray50") # create a layer of borders

mp = 
	ggplot( D_ll ) +   
	mapPacific + 
	geom_point(
		aes(x=`Longitude (dec)`, y=`Latitude (dec)`, col=`Collecting Year`, label1=`Catalog Number` ), alpha = 0.2, size=1
		) +
	scale_color_gradient(low="blue", high="red")

mp
# ggplotly(mp)


```



## U.S. Exploring Expedition 1838-1842

The U.S exploring Expeditions were of historical importance! What specimens in the Peabody IZ Division Collections are from this and where in Pacific are they from?

It turns out that most of these specimens do not have lat and long or precise locality, perhaps to gain this data cross reference with orignial ship logs? Or use the Nearest named place as the locality? 

- Next objective create a map using the nearest named place as the locality

```{r}
D %>% filter(`Collector/Expedition`=="U.S. Exploring Expedition 1838-1842")

```


##U.S Exploring Expedition, Best locations 

finding the best location by filtering for the Expedition name and then selecting the columns in the table that are of interest (Country or nearest locality).  

Not every specimen has both of these variables, so will combine columns to get of both in one column.




```{r}

U.S._ExEx <- D %>% filter(`Collector/Expedition`=="U.S. Exploring Expedition 1838-1842")
US_ExEx_Best_Location <- US_ExEx %>% select('Catalog Number', Country, 'City/Town/Nearest Named Place', 'Latitude (dec)', 'Longitude (dec)')
US_ExEx_Best_Location

US_ExEx_Best_Location$Locality <- paste(US_ExEx_Best_Location$`City/Town/Nearest Named Place`, US_ExEx_Best_Location$Country)

USExEX_Locality <- US_ExEx_Best_Location %>% select('Catalog Number', Locality, 'Latitude (dec)', 'Longitude (dec)')
USExEX_Locality

```

##Specimens by locality

```{r}

Locality = c(table(USExEX_Locality$Locality))
Locality = Locality[order(Locality, decreasing = TRUE)]

Locality_frame = data.frame(Name=names(Locality), N=Locality, stringsAsFactors = FALSE)
rownames(Locality_frame) = NULL
kable(Locality_frame[1:29,] )

```

## Specimens without locality

There are 78 specimens without locality infomation

```{r}

USExEX_Locality %>% filter(Locality=="NA NA")

```


## Specimens by ocean 

goal: to overlay polygons onto map, coloured by density of specimens from each body of water 

```{r}

ocean = c(table(D$`Ocean`))
ocean = ocean[order(ocean, decreasing = TRUE)]

ocean_frame = data.frame(Name=names(ocean), N=ocean, stringsAsFactors = FALSE)
rownames(ocean_frame) = NULL
kable(ocean_frame[1:43,] )



```


#Moving on from specimens by ocean

## Using geospatial tools to create polygone overlaps of specific regions 

What is an ecoregion? 
Bioregionalization of coastal and shelf areas for mapping marine biodiversity 

Original publication 
- https://www.researchgate.net/publication/263705480_Marine_Ecoregions_of_the_World_A_Bioregionalization_of_Coastal_and_Shelf_Areas

Spatial polygon data for plotting
- https://www.worldwildlife.org/publications/marine-ecoregions-of-the-world-a-bioregionalization-of-coastal-and-shelf-areas

```{r}
#for geospatial tools
library(rgdal)
library(maptools)
library(rgeos)

#for some data reduction
library(dplyr)

ogrInfo("/Users/ariannalord/repo/peabody_iz_stats/MEOW", "meow_ecos")

regions <- readOGR("/Users/ariannalord/repo/peabody_iz_stats/MEOW", "meow_ecos")

plot(regions)

provinces <- unionSpatialPolygons(regions, regions$PROVINCE)

#Make a data frame that will have Province level info and above
prov_data <- regions@data %>%
group_by(PROVINCE) %>%
summarise(PROV_CODE = PROV_CODE[1], REALM = REALM[1], RLM_CODE=RLM_CODE[1], Lat_Zone=Lat_Zone[1])

#Merge polygons with new data file
data = data.frame(merge(data.frame(PROVINCE=names(provinces)), prov_data), row.names=row.names(provinces))
provinces <- SpatialPolygonsDataFrame(provinces, data)

#######
#make realms shapefile
########
#make spatial polygons for realms
realms <- unionSpatialPolygons(regions, regions$REALM)

#make new data
realm_data <- regions@data %>%
  group_by(REALM) %>%
  summarise(RLM_CODE = RLM_CODE[1],  Lat_Zone=Lat_Zone[1])

#merge the two!
realms <- SpatialPolygonsDataFrame(realms, 
                                   data=data.frame(
                                     merge(data.frame(REALM=names(realms)),
                                          realm_data),
                                     row.names=row.names(realms)))

##Plot them all
par(mfrow=c(2,2), mar=c(0,0,0,0))
plot(regions, main="Ecoregion", cex.main=5)
plot(provinces, lwd=2, border="red", main="Province")
plot(realms, lwd=2, border="blue", main="Realm")
par(mfrow=c(1,1))

#preparing to plot with color 
regions@data$id = rownames(regions@data)
provinces@data$id = rownames(provinces@data)

#fortify function, this converts a R fucntion into a data frame for ggplot
library(ggplot2)
regions.points = fortify(regions, ECOREGION="id")
provinces.points = fortify(provinces, PROVINCES="id")

#fortified data frames describe the points we want to plot, next we must join the points with the map_data
regions.df = merge(regions.points, regions@data, by="id")
provinces.df = merge(provinces.points, provinces@data, by="id") 

#####Make some ggplots for later visualization
base_ecoregion_ggplot <- ggplot(regions.df) + theme_bw() +
  aes(long,lat,group=group) + 
  geom_polygon(fill=NA) +
  geom_path(color="black") +
  coord_equal() 


base_province_ggplot <- ggplot(provinces.df) + theme_bw() +
  aes(long,lat,group=group) + 
  geom_polygon(fill=NA) +
  geom_path(color="black") +
  coord_equal() 

base_ecoregion_ggplot + ggtitle("Ecoregions")
base_province_ggplot + ggtitle("Provinces")

```


# Now back to the data 
I want to manipulate the specimen data to place the latitude and longitude points into their ecoregions, and then create a choropleth plot dependent on the density of specimens (E.g. as done in the maps in this paper: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3338747/ )

To do this I will create a new table only with the relevant columns and rows.  I will exclude the data that does not have latlong data points and filter by columns of interest (e.g. catalog number, phylum, collector, year)

```{r}
meowrd = D %>% select("Catalog Number", Phylum, "Collector/Expedition", "Collecting Year", "Latitude (dec)", "Longitude (dec)")

meowrd_frame <- na.omit(meowrd)

meowrd_frame

```

#Ecoregions

Now I want to add in the ecoregion data:

combining collections data lat/long and ecoregions data to determine ecoregions of peabody IZ specimens 

```{r}
latlong <- data.frame(Catalog = c(meowrd_frame$"Catalog Number"), lat = c(meowrd_frame$"Latitude (dec)"),                long=c(meowrd_frame$"Longitude (dec)")) 

regionalData <- getRegions(latlong$lat, latlong$long)  
newdata <- cbind(latlong, regionalData)
newdata

```

now, what ecoregions have the most specimens? / great sampling density ?

```{r}

ecoregions_list = c(table(newdata$ECOREGION))
ecoregions_list = ecoregions_list[order(ecoregions_list, decreasing = TRUE)]
ecoregions_frame = data.frame(ECOREGION=names(ecoregions_list), N=ecoregions_list, stringsAsFactors = FALSE)
rownames(ecoregions_frame) = NULL
kable( ecoregions_frame[1:120,] )

```
understanding of distribution 

```{r}

# Simple Bar Plot 
counts <- table(ecoregions_frame$N)
barplot(counts, main="N", 
   xlab="Number of specimens per location")

```




now we make the map expressing this density 

 first is the base province ggplot, next input the values to correspond to colour 
 
 !!!the the hover function aes needs to be fixed here!!
 
 

```{r}

#libraries
library(viridis)
library(plotly)
library(tidyverse)


#breaking up the color to best suite my data distribution 
my_breaks = c(1, 10, 100, 1000, 10000, 20000)

plot.data <- merge(regions.df, ecoregions_frame, by = "ECOREGION")

#plotting color by density of specimens from those locations
p <- ggplot(plot.data, aes(x = long, y = lat, group = group, fill = N)) +
    geom_polygon(
        aes(Ecoregion = 'ECOREGION', "Number of specimens" = 'N'), color = "white", size = 0.2,) +
    coord_equal() +
    scale_fill_viridis(name = "Count", trans = "log",
                      breaks = my_breaks, labels = my_breaks) +
    labs(title = "Number of specimens from Marine Ecoregions of the World",
         fill = NULL) +
    theme_dark() +
    theme(legend.position = "right",
          panel.background = element_rect(fill = NA, colour = "#cccccc"),
          legend.text = element_text(color = "black", size=4))


ggplotly(p)

```
